---
title: "W4exam_MDnotes"
output: html_document
date: "2022-11-02"
---

---
title: "wk4Exam"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(here)
library(sf)
library(tidyr)
library(janitor)

#Load data
Worldshape <- st_read(here("Data","World_Countries__Generalized_.shp" ))
#check CRS
print(Worldshape)

GGIcsv <- read_csv(here::here("Data","HDR21-22_Composite_indices_complete_time_series.csv"))
print(GGIcsv)

```

```{r}
#Checking the class
class(Worldshape)
class(GGIcsv)
```

```{r}
#To check the class of CSV
Datatypelist <- GGIcsv %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_Colnames", 
               values_to="Variable_coltype")
```

```{r}
#cleaning col names of CSV
cleandatacsv <- GGIcsv %>%
  clean_names()
```

```{r}
#selecting required columns
selecteddatacsv <- cleandatacsv%>%
  dplyr::select(contains("iso3"), 
                contains("country"),
                contains("gii_rank_2021"),
                contains("gii_2010"),
                contains("gii_2019"))
```

```{r}

#calculate difference between two

selecteddata <- selecteddatacsv %>% 
  #new column with average of male and female life expectancy
  mutate(Diffinequality= (gii_2019 - gii_2010))%>%
  #select only columns we want
  dplyr::select(iso3,
         country,
         gii_rank_2021, 
         gii_2010,
         gii_2019,
         Diffinequality)

```

```{r}
#slice only the wanted dataset

inequalityIndx<-selecteddata%>% slice(1:195)
print(inequalityIndx)
```

#need to join the countrycodes segment, in csv has iso3c i.e. code uptill three alphabets shp has iso2c i.e. code uptill two alphabets

#so change csv iso3c to iso2c, by creating a new coloumn named iso2c, by selecting country column from ur csv

```{r}
#introduce to iso2c values corresponding to the countryvar in the csv
library(countrycode)

countrycodeiso2<-inequalityIndx%>%
  mutate(iso2c=countrycode(inequalityIndx$country, origin = 'country.name', destination = 'iso2c'))

print(countrycodeiso2)
```

#lets try and join the data based on our iso2c

```{r}
#join data post cleaning

Ljointdata <- Worldshape %>% 
  clean_names() %>%
  left_join(., 
            countrycodeiso2,
            by = c("iso"= "iso2c"))

```

try removing na files and see what happens

```{r}
#removing na files
data<-na.omit(Ljointdata)
```

#make a quick thematic map

```{r}
#map plot
library(tmap)
library(tmaptools)
tmap_mode("plot")

#map of difference in inequality 
qtm(data, 
    fill = "Diffinequality")

```

```{r}
#more plot
tmapworld <- data %>%
  st_bbox(.) %>% 
  tmaptools::read_osm(., type = "osm", zoom = NULL)



tm_shape(data)+
tm_polygons("Diffinequality", 
        style="jenks",
        palette="YlOrBr",
        midpoint=NA,
        title="Difference in inequality",
        alpha = 0.9) + 
  tm_compass(position = c("left", "bottom"),type = "arrow") + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_layout(title = "Global difference in inequality between 2010 and 2019", legend.position = c("right", "bottom"))

```

\`\`\`
